Competing Bandits:\\The Perils of Exploration under Competition

### 2018-11-05 paper/content/introduction.tex
edits

\swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.}  We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage. We find that the data advantage alone still leads to a significant incumbency advantage. Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm. Our results highlight the importance of considering competition dynamics when thinking about the anti-trust question of whether data can serve as a significant barrier to entry in an industry. While data has decreasing returns to scale from a statistical perspective, in a competitive environment the relative differences in the data possessed by the firms may lead to a large competitive benefit due to the feedback loops discussed previously. In our model this effect is amplified when the incumbent commits to a more advanced bandit algorithm since, in the incumbency period, the firm gathers data on possible actions more efficiently leading to the relative gap upon entry to be larger. This is important when exploration can have larger reputational consequences.}\swcomment{guy, could you make another pass over above}

@@

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    (see \cite{DynAuctions-survey10} for background),
online ad auctions, \eg
    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm,Amin-auctions-nips13},
and human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.\swcomment{not sure we need to talk about dueling bandits}
\gaedit{There is a superficial similarity (in name only) between this paper and the line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}). The latter is not about competing
bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".

The study of whether data can give a significant \swdelete{competitive} advantage
has been discussed theoretically in \cite{varian2018artificial,
  lambrecht2015can} and empirically in
\cite{bajari2018impact}. \swedit{While they argue that small amounts
  of additional data do not provide significant improvement in
  training a more predictive model, this set of results focus on learning in isolation, which overlooks certain aspects in a competition.}
\swdelete{
These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.} We ask how additional data helps while firms simultaneously solve a machine learning problem and compete for consumers. In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm. Additionally, the first-mover advantage has been well-studied in economics and marketing \cite{kerin1992first}. }

>>>>

\swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.}  We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage. We find that the data advantage alone still leads to a significant incumbency advantage. Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm. Our results highlight the importance of considering competition dynamics when thinking about the anti-trust question of whether data can become a significant barrier to entry in an industry. While data has decreasing returns to scale from a statistical perspective, in a competitive environment the relative differences in the data possessed by the firms may lead to a large competitive benefit due to the feedback loops discussed previously. In our model this effect is amplified when the incumbent commits to a more advanced bandit algorithm since, in the incumbency period, the firm gathers data on possible actions more efficiently leading to the relative gap upon entry to be larger. This is important when exploration can have larger reputational consequences.}

@@

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    (see \cite{DynAuctions-survey10} for background),
online ad auctions, \eg
    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm,Amin-auctions-nips13},
and human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
\swedit{There is also a line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}) that also studies a form of competition with MAB problems. However, their framework is fundamentally different from ours in that they study learning from setting ``duels'' between pairs of arms, and we study how learning algorithms compete with each other for data.
}

The study of whether data can give a significant \swdelete{competitive} advantage
has been discussed theoretically in \cite{varian2018artificial,
  lambrecht2015can} and empirically in
\cite{bajari2018impact}. \swedit{While they argue that small amounts
  of additional data do not provide significant improvement in
  training a more predictive model, this set of results focus on learning in isolation, which overlooks certain aspects in a competition.
\swdelete{
These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.} We ask how additional data helps while firms simultaneously solve a machine learning problem and compete for consumers. In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm. Additionally, the first-mover advantage has been well-studied in economics and marketing \cite{kerin1992first}. }
